Categories of Optics

### 2018-06-09 paper.tex
Some typos/corrections

\documentclass[11pt,a4paper]{article}
\usepackage{authblk}
%\usepackage[utf8]{inputenc}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\let\amssquare\square
\usepackage{mathtools}
%\usepackage{stmaryrd}
%\usepackage{tensor}
\usepackage[mathscr]{eucal}
\usepackage{url}
%\usepackage{marvosym}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}

@@

\begin{remark}
Note that the definition of the $\Optic$ category involves a coend possibly indexed by a large category. To be careful, we should only ever discuss $\Optic$ categories where we know that the coend exists by some other means, e.g., by exhibiting an isomorphism with some other set. For many---but not all---of the examples given later, we provide such an isomorphism.

@@

This preserves identities, as the identity on an object $(S, S')$ in $\Optic_\M$ is defined to be exactly $\langle \lambda^{-1}_S, \lambda_{S'} \rangle$.

To check functoriality, suppose we have $(f, g) : (S, S') \to (A, A')$ and $(f', g') : (A, A') \to (B, B')$ in $\C \times \C^\op$. Then:
\begin{align*}
\iota(f', g') \circ \iota(f, g) 
&= \langle \lambda^{-1}_B f', g' \lambda_{B'} \rangle \circ \langle \lambda^{-1}_A f, g \lambda_{A'} \rangle \\
&= \langle (I\otimes (\lambda^{-1}_B f'))\lambda^{-1}_A f, g \lambda_{A'} (I\otimes (g' \lambda_{B'}))\rangle && \text{(By definition of $\circ$)}\\
&= \langle (I \otimes \lambda^{-1}_B) (I \otimes f')\lambda^{-1}_A f, g \lambda_{A'} (I \otimes g')(I\otimes \lambda_{B'})\rangle && \text{(Functoriality of $I \otimes -$)}\\
&= \langle (I\otimes \lambda^{-1}_B) \lambda^{-1}_B f' f, g g' \lambda_{B'} (I\otimes \lambda_{B'})\rangle && \text{(Naturality of $\lambda$)}\\
&= \langle (\lambda^{-1}_I \otimes B) \lambda^{-1}_B f' f, g g' \lambda_{B'} (\lambda_I \otimes B')\rangle && \text{(Unitality of action)} \\
&= \langle \lambda^{-1}_B f' f, g g' \lambda_{B'} (\lambda_I \otimes B') (\lambda^{-1}_I \otimes B') \rangle && \text{(Coend relation)}  \\
&= \langle \lambda^{-1}_B f'f, g g' \lambda_{B'} \rangle \\
&= \iota(f'f, gg')
\end{align*}
Graphically, there is not much to do:
\begin{center}
\input{diagrams/iota-functorial-left.tikz}
\qquad \raisebox{0.3cm}{$=$} \qquad 
\input{diagrams/iota-functorial-right.tikz}
\end{center}

@@

\begin{proposition}
$\iota : \C \times \C^\op \to \Optic_\C$ is a strong monoidal functor. 
\end{proposition}
\begin{proof}
\todo{todo}
\end{proof}

@@

\begin{proposition}
\todo{Entrywise Coproduct}
\end{proposition}

\begin{remark}
The motivating example here is that of coproducts distributing over the tensor product.
\end{remark}

@@

\begin{proposition}
\label{prop-optic-functor}
Collectively this operation defines a \todo{pseudo?}functor \[\Optic : \SymmMonCat \to \SymmMonCat.\] \qed
\end{proposition}

@@

We assume temporarily that $\C$ is a \emph{strict} symmetric monoidal category. It will be convenient to equip $\Optic_\C$ with a slightly different symmetric monoidal structure: \todo{And the universal property given in this section is an argument for this being the ``morally correct'' tensor}

@@

@@

Let $\inside : \Optic_\M(S, A) \to \int^{M \in \M} \C(M A, M A)$ be the map induced by $\inside(\langle l, r \rangle) = \langle r ; l \rangle$. We might ask that instead of requiring $r;l = \phi A$ exactly, we have $\langle r ; l \rangle = \langle \phi A \rangle$ in $\int^{M \in \M} \C(M A, M A)$2  . In fact, this is equivalent:

@@

We now have to slog through evaluating $\Phi(\once(p))$ and $\Phi(\twice(p))$.
\begin{alignat*}{3}
\Phi(\once(\langle [\id_S, \fget],  \fput \rangle)) &= 
\Phi(&&\langle [\id_S, \fget], \id_{S \times A}, \fput \rangle) \\
&= (&&\pi_2 [\id_S, \fget], \\
&&&\pi_2 \id_{S \times A} (\pi_1 [\id_S, \fget] \times A), \\
&&& \fput (\pi_1 \id_{S \times A} (\pi_1 [\id_S, \fget] \times A) \times A) \quad) \\
%
&= (&&\fget, \\
&&& \pi_2, \\
&&& \fput \pi_{1,3} \quad) \\
%%%%
%%%%
\Phi(\twice(\langle [\id_S, \fget],  \fput \rangle)) &= 
\Phi(&&\langle [\id_S, \fget], [\id_S, \fget]\fput, \fput \rangle) \\
&= (&&\pi_2 [\id_S, \fget], \\
&&&\pi_2 [\id_S, \fget]\fput (\pi_1 [\id_S, \fget] \times A), \\
&&& \fput (\pi_1 [\id_S, \fget]\fput (\pi_1 [\id_S, \fget] \times A) \times A)) \quad) \\
%
&= (&&\fget, \\
&&& \fget \; \fput (\id_S \times A), \\
&&& \fput (\fput (\id_S \times A) \times A) \quad) \\
%
&= (&&\fget, \\
&&& \fget \; \fput, \\
&&& \fput (\fput \times A) \quad)
\end{alignat*}
So $\Phi(\once(p))$ being equal to $\Phi(\twice(p))$ is exactly equivalent to the $\fput\fget$ and $\fput\fput$ laws holding.
\end{proof}

@@

\begin{proposition}
When $\C = \Set$\todo{, and probably more generally,} the third law is implied by the other two.
\end{proposition}
\begin{proof}
The key is that in $\Set$, for any map $f : X \to Y$, the set $Y$ is equal to the union of $\im f$ and its complement. The first law implies that $\freview$ is injective, so $S \cong C \sqcup A$ for some complement $C$. Identifying $A$ with its image in $S$, the second law implies that if $a\in A \subset S$ then $\fmatching(a) = \inr(a)$ and if $c\in C \subset S$ then $\fmatching(c) = \inl(c)$. The third law can then be verified pointwise by checking both cases $a\in A \subset S$ and $c\in C \subset S$ separately.
\end{proof}

@@

\begin{proposition}
An iso $\langle l, r \rangle$ is lawful iff (as expected) $l$ and $r$ are mutual inverses.
\end{proposition}
\begin{proof}
\[ \int^{M_1, M_2 \in \M} \C(S, M_1 \otimes A) \times \C(M_1 \otimes A, M_2 \otimes A) \times \C(M_2 \otimes A, S) \]
specialises in this case to just
\[ \C(S, A) \times \C(A, A) \times \C(A, S) \]
The condition $\outside(\langle l, r \rangle) = \id_S$ is the claim that $rl = \id_S$, and $\once(\langle l, r \rangle) = (l, \id_A, r)$ is equal to $\twice(\langle l, r \rangle) = (l, lr, r)$ iff $lr = \id_A$.s
\end{proof}

@@

\subsection{Polymorphic Optics}
Haskell's optics \todo{find an older reference from DB theory?} allow \emph{polymorphic updates}, where the type of the codomain of the lens can be changed by an update, causing a corresponding change in the type of the domain. As an example, the lens into the first entry of a tuple has type
\begin{center}
\texttt{first :: Lens (a, b) (a', b) a a'},
\end{center}
\todo{Why is the spacing strange above?} then:
\begin{center}
\texttt{set first (1, 5) "hello" == ("hello", 5)}
\end{center}
Note that the type has been changed from $\texttt{(Int, Int)}$ to $\texttt{(String, Int)}$.

@@

\subsection{Optics}
\todo{Figure out the most sensible order to prove these things in}

\begin{definition}
For an object $A \in \C$, the \emph{exchange profunctor} $E_A$ is defined to be $\C(-, A) \times \C(A, {=})$.
\end{definition}

@@

\begin{proposition}[Profunctor Optics are Optics]
\label{prop-profunctor-optics-are-optics}
\begin{align*}
[\Tamb(\M), \Set]((U-)(A,A),(U-)(S,S)) &\cong \Optic_\M(S, A)
\end{align*}
\end{proposition}
\begin{proof}
We have the chain of isomorphisms:
\begin{align*}
&[\Tamb(\M), \Set]((U-)(A,A),(U-)(S,S)) \\
\cong \;&[\Tamb(\M), \Set](\Tamb(\M)(\Pastro_\M E_A, -), (U-)(S,S))  && \text{(by Lemma \ref{lemma-rep})}\\
\cong \;&(U\Pastro_\M E_A)(S,S)  && \text{(by Yoneda)} \\
= \;&\int^{M \in \M} \int^{C,D \in \C} \C(S, MC) \times E_A(C,D) \times \C(MD,S) \\
= \;&\int^{M \in \M} \int^{C,D \in \C} \C(S, MC) \times \C(C,A) \times \C(A,D) \times \C(MD,S) \\
\cong \;&\int^{M \in \M} \C(S, MA) \times \C(MA, S)  && \text{(by co-Yoneda twice)} \\
= \;&\Optic_\M(S, A)
\end{align*}
\todo{This works for $\Optic_\M((S, S'), (A, A'))$ just fine, just define $E_{A,A'}$ in the obvious way.}

\todo{Flip the order of the isomorphisms top to bottom?}
\end{proof}

For $p : S \hto A$, let $\tilde{p} : (U-)(A,A) \Rightarrow (U-)(S,S)$ denote the corresponding natural transformation under this isomorphism, and for $t : (U-)(A,A) \Rightarrow (U-)(S,S)$, let $\hat{t} : S \hto A$ be the corresponding optic.

\begin{corollary}
A profunctor optic is determined by its component at $\Pastro_\M E_A$, and furthermore, this component is determined by its value on $\langle \id_A, \id_A, \id_A, \id_A \rangle \in (U \Pastro_\M E_A)(A, A)$.
\end{corollary}
\begin{proof}
This is the content of the first two isomorphisms above.
\todo{fix notation in the following}
Explicitly, let $p_{\Pastro_\M E_A}(\langle \id_A, \id_A, \id_A, \id_A \rangle) = \langle l, \id_A, \id_A, r \rangle$, where $l : S \to M A$ and $r : M A \to S$. Then the component of $p$ at any other Tambara module $P$ is given by
\begin{align*}
p_P = (UP)(l,r) \alpha_{A,A,M}
\end{align*}
where $\alpha$ is the module structure for $P$.
\end{proof}

There is yet another isomorphic description of optics: \todo{Reference proof in PS}

\begin{proposition}
$\Optic_\M(S, A)$ is isomorphic to $\Prof(E_S, U \Phi_\M E_A)$. \todo{Equivalently to $\Tamb(\M)(\Phi_\M E_S, \Phi_\M E_A)$ by adjointness}
\end{proposition}
\begin{proof}
\begin{align*}
&\Prof(E_S, U \Phi_\M E_A) \\
\cong\;&\int_{X,Y \in \C} \Set(E_S(X,Y), (U \Phi_\M E_A)(X,Y)) \\
= \;&\int_{X,Y \in \C} \Set(\C(X,A) \times \C(A,Y), (U \Phi_\M E_A)(X,Y)) \\
\cong \;& (U \Phi_\M E_A)(S,S) \\
= \;&\int^{M \in \M} \C(S, MA) \times \C(MA, S) \\
= \;&\Optic_\M(S, A)
\end{align*}
\end{proof}

\todo{This gives one direction for generalising optics: are there other easy to describe comonoid objects in $\Prof$?}

@@

The next question is: can we characterise the profunctor optics that correspond to lawful optics? The plan is to chase through the above isomorphisms to see where the lawful optics end up.

The exchange profunctor has a canonical comonoid structure. The comultiplication $\Delta : E_A \to E_A \odot E_A$ is given by the composite
\begin{align*}
&E_A(X,Y) \\
= \quad& \C(X, A) \times \C(A, Y) \\
\xrightarrow{[\pi_1, \const_{\id_A}, \const_{\id_A}, \pi_2]} \quad& \C(X, A) \times \C(A, A) \times \C(A, A) \times \C(A, Y) \\
\xrightarrow{\copr_A} \quad& \int^{C \in \C} \C(X, A) \times \C(A, C) \times \C(C, A) \times \C(A, Y) \\
= \quad& \int^{C \in \C} E_A(X,C) \times E_A(C,Y) \\
= \quad&  (E_A \odot E_A)(X,Y)
\end{align*}
and the counit $\varepsilon : E_A \to I$ simply composes the two morphisms. Because $\Pastro_\M$ is oplax monoidal, the Tambara module $\Pastro_\M E_A$ has an induced comonoid structure.

\begin{theorem}
An optic $p : S \hto A$ is lawful iff the corresponding natural transformation $E_S \Rightarrow U \Phi_\M E_A$ is a comonoid morphism.
\end{theorem}
\begin{proof}

\begin{lemma}
For an object $X$ in a monoidal category $\C$, a comonoid structure $(X,\Delta,\varepsilon)$ is equivalent to a lax monoidal structure on $\C(X, -)$.
\end{lemma}
\begin{proof}
\todo{not hard}
\end{proof}

Therefore the functor $\Tamb(\M)(\Pastro_\M E_A, -)$ is canonically lax monoidal, and so the isomorphic functor $(U-)(A,A)$ is also.

\todo{Once this is proven, is the remainder of the section for free?}

\begin{proposition}
If $p : S \hto A$ is a lawful optic then the associated natural transformation $\tilde{p} : (U-)(A,A) \to (U-)(S,S)$ is lax monoidal.
\end{proposition}
\begin{proof}
Let $\langle l, r \rangle$ be a representative for $p$ in $\Optic_\M(S,A)$ via the isomorphism of Proposition \ref{prop-profunctor-optics-are-optics}, so that
\begin{align*}
\tilde{p}_P = (UP)(l,r) \alpha_{A,A,M}
\end{align*}
on each Tambara module $P$.

To show $\tilde{p}$ is monoidal, we must show that both
\[
\begin{tikzcd}
UP(A,A) \times UQ(A,A) \ar[r, "\tilde{p}_P \times \tilde{p}_Q"] \ar[d, swap] & UP(S, S) \times UQ(S, S)  \ar[d] \\
U(P \odot Q)(A,A)  \ar[r, "\tilde{p}_{P \odot Q}", swap] & U(P \odot Q)(S,S)
\end{tikzcd}
\]
and
\[
\begin{tikzcd}
1 \ar[d] \ar[dr] & \\
UI(A,A)  \ar[r, "\tilde{p}_I", swap] & UI(S,S)
\end{tikzcd}
\]
commute. For the first we have
\begin{align*}
\langle (\tilde{p}_P \times \tilde{p}_Q)(x, y) \rangle &= \langle (UP)(l,r) \alpha_{A,A,M} x,  (UQ)(l,r) \beta_{A,A,M} y\rangle \\
&= \langle (UP)(l,lr) \alpha_{A,A,M} x, (UQ)(\id_{MA},r) \beta_{A,A,M} y\rangle && \text{(by the coend relations)} \\
&= \langle (UP)(l,\phi A) \alpha_{A,A,M} x, (UQ)(\id_{MA},r) \beta_{A,A,M} y\rangle && \text{(as $lr = \phi A$)} \\
&= \langle (UP)((\phi A)l,\id_{MA}) \alpha_{A,A,M} x, (UQ)(\id_{MA},r) \beta_{A,A,M} y\rangle && \text{(by the coend relations again)} \\
&= \langle (UP)(lrl,\id_{MA}) \alpha_{A,A,M} x, (UQ)(\id_{MA},r) \beta_{A,A,M} y\rangle && \text{(as $lr = \phi A$ again)} \\
&= \langle (UP)(l,\id_{MA}) \alpha_{A,A,M} x, (UQ)(\id_{MA},r) \beta_{A,A,M} y\rangle && \text{(as $rl = \id_S$)} \\
&= (U(P \odot Q))(l,r) \langle \alpha_{A,A,M} x, \beta_{A,A,M} y \rangle && \text{(by definition of $P \odot Q$)} \\
&= (U(P \odot Q))(l,r) (\alpha_{A,A,M} \odot \beta_{A,A,M}) \langle x, y \rangle && \text{(by definition of $\alpha \odot \beta$)} \\
&= \tilde{p}_{P \odot Q} \langle x, y \rangle
\end{align*}
And for the second:
\begin{align*}
\tilde{p}_I(\id_A) = (UI)(l,r) (M \id_A) = r \id_{MA} l = rl = \id_S
\end{align*}
as required.
\end{proof}

%\begin{proposition}
%Let $\C = \Set$ and suppose $p : S \hto A$ is a lens such that the associated natural transformation $p : (U-)(A,A) \to (U-)(S,S)$ is lax monoidal. Then $p$ is lawful.
%\end{proposition}
%\begin{proof}
%\todo{Show that it satisfies the getput, putget and putput laws. This strategy works only with $\times$ and in $\Set$.}
%\end{proof}

\todo{Here are some thoughts:}
The converse is more difficult, and unfortunately requires another fiddly proof by induction on the coend relation. The idea is that any profunctor optic is characterised by its component at $\Pastro_\M E_A$, so the result should follow from the commutativity of the square for lax monoidality when $P = Q = \Pastro_\M E_A$.

For the remainder of the section, fix a lax monoidal natural transformation $t : (U-)(A,A) \to (U-)(S,S)$ and suppose $\hat{t} = \langle l, r \rangle$ for $l : S \to MA$ and $r : MA \to S$. Commutativity of the triangle implies immediately that $rl = \id_S$.

>>>>

\documentclass[11pt,a4paper]{article}
\usepackage{authblk}
%\usepackage[utf8]{inputenc}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\let\amssquare\square
\usepackage{mathtools}
%\usepackage{stmaryrd}
%\usepackage{tensor}
\usepackage[mathscr]{eucal}
\usepackage{url}
%\usepackage{marvosym}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\usepackage{minted}

@@

\begin{remark}
Note that the homsets of $\Optic_\C$ are given by a coend indexed by a possibly large category, so their existence is not guaranteed by the cocompleteness of $\Set$. We should be careful to only discuss $\Optic$ categories where we know that the coends exist by some other means, e.g., by exhibiting an isomorphism with some other set. For many---but not all---of the examples given later, we provide such an isomorphism.

@@

This preserves identities, as the identity on an object $(S, S')$ in $\Optic_\M$ is defined to be exactly $\langle \lambda^{-1}_S, \lambda_{S'} \rangle$. To check functoriality, suppose we have $(f, g) : (S, S') \to (A, A')$ and $(f', g') : (A, A') \to (B, B')$ in $\C \times \C^\op$. Then:
\begin{align*}
\iota(f', g') \circ \iota(f, g) 
&= \langle \lambda^{-1}_B f', g' \lambda_{B'} \rangle \circ \langle \lambda^{-1}_A f, g \lambda_{A'} \rangle \\
&= \langle (I\otimes (\lambda^{-1}_B f'))\lambda^{-1}_A f, g \lambda_{A'} (I\otimes (g' \lambda_{B'}))\rangle && \text{(By definition of $\circ$)}\\
&= \langle (I \otimes \lambda^{-1}_B) (I \otimes f')\lambda^{-1}_A f, g \lambda_{A'} (I \otimes g')(I\otimes \lambda_{B'})\rangle && \text{(Functoriality of $I \otimes -$)}\\
&= \langle (I\otimes \lambda^{-1}_B) \lambda^{-1}_B f' f, g g' \lambda_{B'} (I\otimes \lambda_{B'})\rangle && \text{(Naturality of $\lambda$)}\\
&= \langle (\lambda^{-1}_I \otimes B) \lambda^{-1}_B f' f, g g' \lambda_{B'} (\lambda_I \otimes B')\rangle && \text{(Unitality of action)} \\
&= \langle \lambda^{-1}_B f' f, g g' \lambda_{B'} (\lambda_I \otimes B') (\lambda^{-1}_I \otimes B') \rangle && \text{(Coend relation)}  \\
&= \langle \lambda^{-1}_B f'f, g g' \lambda_{B'} \rangle \\
&= \iota(f'f, gg')
\end{align*}
Graphically, there is not much to do:
\begin{center}
\input{diagrams/iota-functorial-left.tikz}
\qquad \raisebox{0.3cm}{$=$} \qquad 
\input{diagrams/iota-functorial-right.tikz}
\end{center}

@@

\begin{proposition}
$\iota : \C \times \C^\op \to \Optic_\C$ is a monoidal functor. 
\end{proposition}
\begin{proof}
This is clear, as the structure morphisms in $\Optic_\C$ are exactly the structure morphisms in $\C \times \C^\op$ under $\iota$.
\end{proof}

@@

%\begin{proposition}
%\todo{Entrywise Coproduct}
%\end{proposition}
%
%\begin{remark}
%The motivating example here is that of coproducts distributing over the tensor product.
%\end{remark}

@@

\begin{proposition}
\label{prop-optic-functor}
Collectively this operation defines a 2-functor \[\Optic : \SymmMonCat \to \SymmMonCat.\] \qed
\end{proposition}
\todo{Haven't actually checked natural transformations}
\todo{Also we run into the large coend issue again, it's probably not well defined on every symm mon cat.}

@@

We assume temporarily that $\C$ is a \emph{strict} symmetric monoidal category. It will be convenient to equip $\Optic_\C$ with a slightly different symmetric monoidal structure. The universal property for $\Optic_\C$ given in this section is an argument for this being the ``morally correct'' tensor, although it does seem a little strange.

@@

\begin{remark}
Jeremy Gibbons notes (originally in the case of ordinary lenses) that the first law is equivalent to requiring that $p$ composed with the connector $c_A$ is equal to the connector $c_S$. This is an appealing description from a string diagram standpoint, but I do not know if there is a similar description for the second law. \todo{Something to do with embedding $\Optic_\C$ into a compact closed category then using the unit to compose the lenses vertically? Should I draw a picture of this?}
\end{remark}

@@

Let $\inside : \Optic_\M(S, A) \to \int^{M \in \M} \C(M A, M A)$ be the map induced by $\inside(\langle l, r \rangle) = \langle r ; l \rangle$. We might ask that instead of requiring $r;l = \phi A$ exactly, we have $\langle r ; l \rangle = \langle \phi A \rangle$ in $\int^{M \in \M} \C(M A, M A)$. In fact, this is equivalent:

@@

We now have to slog through evaluating $\Phi(\once(p))$ and $\Phi(\twice(p))$.
\begin{alignat*}{3}
\Phi(\once(\langle [\id_S, \fget],  \fput \rangle)) &= 
\Phi(&&\langle [\id_S, \fget], \id_{S \times A}, \fput \rangle) \\
&= (&&\pi_2 [\id_S, \fget], \\
&&&\pi_2 \id_{S \times A} (\pi_1 [\id_S, \fget] \times A), \\
&&& \fput (\pi_1 \id_{S \times A} (\pi_1 [\id_S, \fget] \times A) \times A) \quad) \\
%
&= (&&\fget, \\
&&& \pi_2, \\
&&& \fput \pi_{1,3} \quad) \\
%%%%
%%%%
\Phi(\twice(\langle [\id_S, \fget],  \fput \rangle)) &= 
\Phi(&&\langle [\id_S, \fget], [\id_S, \fget]\fput, \fput \rangle) \\
&= (&&\pi_2 [\id_S, \fget], \\
&&&\pi_2 [\id_S, \fget]\fput (\pi_1 [\id_S, \fget] \times A), \\
&&& \fput (\pi_1 [\id_S, \fget]\fput (\pi_1 [\id_S, \fget] \times A) \times A)) \quad) \\
%
&= (&&\fget, \\
&&& \fget \; \fput (\id_S \times A), \\
&&& \fput (\fput (\id_S \times A) \times A) \quad) \\
%
&= (&&\fget, \\
&&& \fget \; \fput, \\
&&& \fput (\fput \times A) \quad)
\end{alignat*}
So checking componentwise, $\Phi(\once(p))$ being equal to $\Phi(\twice(p))$ is exactly equivalent to the $\fput\fget$ and $\fput\fput$ laws holding.
\end{proof}

@@

\begin{proposition}
When $\C = \Set$, the third law is implied by the other two.
\end{proposition}
\begin{proof}
The key is that in $\Set$, for any map $f : X \to Y$, the set $Y$ is equal to the union of $\im f$ and its complement. The first law implies that $\freview$ is injective, so $S \cong C \sqcup A$ for some complement $C$. Identifying $A$ with its image in $S$, the second law implies that if $a\in A \subset S$ then $\fmatching(a) = \inr(a)$ and if $c\in C \subset S$ then $\fmatching(c) = \inl(c)$. The third law can then be verified pointwise by checking both cases $a\in A \subset S$ and $c\in C \subset S$ separately.
\end{proof}

@@

\begin{proposition}
An iso $\langle l, r \rangle$ is lawful iff (as expected) $l$ and $r$ are mutual inverses.
\end{proposition}
\begin{proof}
\[ \int^{M_1, M_2 \in \M} \C(S, M_1 \otimes A) \times \C(M_1 \otimes A, M_2 \otimes A) \times \C(M_2 \otimes A, S) \]
specialises in this case to just
\[ \C(S, A) \times \C(A, A) \times \C(A, S) \]
The condition $\outside(\langle l, r \rangle) = \id_S$ is the claim that $rl = \id_S$, and $\once(\langle l, r \rangle) = (l, \id_A, r)$ is equal to $\twice(\langle l, r \rangle) = (l, lr, r)$ iff $lr = \id_A$.
\end{proof}

@@

\subsection{Polymorphic Optics}
Haskell's optics \todo{find an older reference from DB theory?} allow \emph{polymorphic updates}, where the type of the codomain of the lens can be changed by an update, causing a corresponding change in the type of the domain. As an example, the lens into the first entry of a tuple has type
\begin{minted}{haskell}
first :: Lens (a, b) (a', b) a a'
\end{minted}
then:
\begin{minted}{haskell}
set first (1, 5) "hello" == ("hello", 5)
\end{minted}
Note that the type has been changed from $\texttt{(Int, Int)}$ to $\texttt{(String, Int)}$.

@@

\subsection{Optics}
\begin{definition}
For an object $A \in \C$, the \emph{exchange profunctor} $E_A$ is defined to be $\C(-, A) \times \C(A, {=})$.
\end{definition}

@@

\begin{proposition}[Profunctor Optics are Optics]
\label{prop-profunctor-optics-are-optics}
\begin{align*}
[\Tamb(\M), \Set]((U-)(A,A),(U-)(S,S)) &\cong \Optic_\M(S, A)
\end{align*}
\end{proposition}
\begin{proof}
We have the chain of isomorphisms:
\begin{align*}
&[\Tamb(\M), \Set]((U-)(A,A),(U-)(S,S)) \\
\cong \;&[\Tamb(\M), \Set](\Tamb(\M)(\Pastro_\M E_A, -), (U-)(S,S))  && \text{(by Lemma \ref{lemma-rep})}\\
\cong \;&(U\Pastro_\M E_A)(S,S)  && \text{(by Yoneda)} \\
= \;&\int^{M \in \M} \int^{C,D \in \C} \C(S, MC) \times E_A(C,D) \times \C(MD,S) \\
= \;&\int^{M \in \M} \int^{C,D \in \C} \C(S, MC) \times \C(C,A) \times \C(A,D) \times \C(MD,S) \\
\cong \;&\int^{M \in \M} \C(S, MA) \times \C(MA, S)  && \text{(by co-Yoneda twice)} \\
= \;&\Optic_\M(S, A)
\end{align*}
\end{proof}

The same proof works in the more general case of $\Optic_\M((S, S'), (A,A'))$, defining $E_{A, A'}$ to be $\C(-, A) \times \C(A', {=})$.

For $p : S \hto A$, let $\tilde{p} : (U-)(A,A) \rightarrow (U-)(S,S)$ denote the corresponding natural transformation under this isomorphism, and for $t : (U-)(A,A) \rightarrow (U-)(S,S)$, let $\hat{t} : S \hto A$ be the corresponding optic.

\begin{corollary}
A profunctor optic $p$ is determined by its component at $\Pastro_\M E_A$, and furthermore, this component is determined by its value on $\langle \id_A, \id_A, \id_A, \id_A \rangle \in (U \Pastro_\M E_A)(A, A)$.
\end{corollary}
\begin{proof}
This is the content of the first two isomorphisms above. Explicitly, suppose $\hat{p} = \langle l, r \rangle$ with $l : S \to MA$ and $r : MA \to S$. Then for any Tambara module $P$, the component of $p$ at $P$ is 
\begin{align*}
p_P = (UP)(l,r) \alpha_{A,A,M}
\end{align*}
where $\alpha$ is the module structure for $P$. In particular, 
\[
p_{\Pastro_\M E_A}(\langle \id_A, \id_A, \id_A, \id_A \rangle) = \langle l, \id_A, \id_A, r \rangle
\]
\end{proof}

There is yet another isomorphic description of optics, which may be an unexplored avenue for generalisation.

\begin{proposition}
$\Optic_\M(S, A)$ is isomorphic to $\Prof(E_S, U \Phi_\M E_A)$.
\end{proposition}
\begin{proof}
\begin{align*}
&\Prof(E_S, U \Phi_\M E_A) \\
\cong\;&\int_{X,Y \in \C} \Set(E_S(X,Y), (U \Phi_\M E_A)(X,Y)) \\
= \;&\int_{X,Y \in \C} \Set(\C(X,A) \times \C(A,Y), (U \Phi_\M E_A)(X,Y)) \\
\cong \;& (U \Phi_\M E_A)(S,S) \\
= \;&\int^{M \in \M} \C(S, MA) \times \C(MA, S) \\
= \;&\Optic_\M(S, A)
\end{align*}
\end{proof}

@@

The next goal is to characterise the profunctor optics that correspond to lawful optics.

The exchange profunctor has a canonical comonoid structure, where the comultiplication $\Delta : E_A \to E_A \odot E_A$ and counit $\varepsilon : E_A(X, Y) \to \C(X, Y)$ are given by
\begin{align*}
\Delta((f, g)) &= \langle f, \id_A, \id_A, g \rangle \\
\varepsilon((f, g)) &= gf
\end{align*}
Because $\Pastro_\M$ is oplax monoidal, the profunctor $U\Pastro_\M E_A$ has an induced comonoid structure, given by
\begin{align*}
\Delta(\langle l, r \rangle) &= \langle l, lr, r \rangle \\
\varepsilon(\langle l, r \rangle) &= rl
\end{align*}

Comparing these definitions with the definition of a lawful optic, we have immediately:

\begin{proposition}
An optic $p : S \hto A$ is lawful iff the corresponding natural transformation $E_S \rightarrow U \Phi_\M E_A$ is a comonoid morphism. \qed
\end{proposition}

All that is needed to complete the connection with profunctor optics is the following standard result in category theory.

\begin{lemma}
For an object $X$ in a monoidal category $(\C, \otimes, I)$, a comonoid structure $(X,\Delta,\varepsilon)$ is equivalent to a lax monoidal structure on $\C(X, -)$. Further, comonoid morphisms $(X_1,\Delta_1,\varepsilon_1) \to (X_2,\Delta_2,\varepsilon_2)$ are equivalent to monoidal natural transformations $\C(X, -) \to \C(Y, -)$.
\end{lemma}
\begin{proof}
For a comonoid $(X,\Delta,\varepsilon)$, define a lax monoidal structure 
\begin{align*}
\phi &: 1 \to \C(X, I) \\
\phi_{A, B} &: \C(X, A) \times \C(X, B) \to \C(X, A \otimes B) 
\end{align*}
by $\phi = \varepsilon$ and $\phi_{A, B}(f, g) = (f \otimes g) \Delta$.

From such a lax monoidal structure, we recover the comonoid maps by $\varepsilon = \phi$ and $\Delta = \phi_{X, X}(\id_X, \id_X)$.

\todo{check equations, and the rest}
\end{proof}

\begin{theorem}
$p : S \hto A$ is a lawful optic iff the associated natural transformation $\tilde{p} : (U-)(A,A) \to (U-)(S,S)$ is monoidal with respect to the canonical lax monoidal structures on $(U-)(A,A)$ and $(U-)(S,S)$.
\end{theorem}
\begin{proof}
$p : S \hto A$ is lawful iff $E_S \rightarrow U \Phi_\M E_A$ is a comonoid morphism, iff $\Phi_\M E_S \rightarrow \Phi_\M E_A$ is a comonoid morphism by adjointness. Now apply the previous lemma.
\end{proof}

\begin{remark}
For Haskell's profunctor lenses, these laws take the form:
\begin{minted}{haskell}
l :: Tambara m p => p a a -> p s s

l id == id
l (Procompose p q) == Procompose (l p) (l q)
\end{minted}
\end{remark}

%\begin{proposition}
%If $p : S \hto A$ is a lawful optic then the associated natural transformation $\tilde{p} : (U-)(A,A) \to (U-)(S,S)$ is lax monoidal.
%\end{proposition}
%Let $\langle l, r \rangle$ be a representative for $p$ in $\Optic_\M(S,A)$ via the isomorphism of Proposition \ref{prop-profunctor-optics-are-optics}, so that
%\tilde{p}_P = (UP)(l,r) \alpha_{A,A,M}
%on each Tambara module $P$.
%To show $\tilde{p}$ is monoidal, we must show that both
%\[
%\begin{tikzcd}
%UP(A,A) \times UQ(A,A) \ar[r, "\tilde{p}_P \times \tilde{p}_Q"] \ar[d, swap] & UP(S, S) \times UQ(S, S)  \ar[d] \\
%U(P \odot Q)(A,A)  \ar[r, "\tilde{p}_{P \odot Q}", swap] & U(P \odot Q)(S,S)
%\end{tikzcd}
%\]
%and
%\[
%\begin{tikzcd}
%1 \ar[d] \ar[dr] & \\
%UI(A,A)  \ar[r, "\tilde{p}_I", swap] & UI(S,S)
%\end{tikzcd}
%\]
%commute. For the first we have
%\langle (\tilde{p}_P \times \tilde{p}_Q)(x, y) \rangle &= \langle (UP)(l,r) \alpha_{A,A,M} x,  (UQ)(l,r) \beta_{A,A,M} y\rangle \\
%&= \langle (UP)(l,lr) \alpha_{A,A,M} x, (UQ)(\id_{MA},r) \beta_{A,A,M} y\rangle && \text{(by the coend relations)} \\
%&= \langle (UP)(l,\phi A) \alpha_{A,A,M} x, (UQ)(\id_{MA},r) \beta_{A,A,M} y\rangle && \text{(as $lr = \phi A$)} \\
%&= \langle (UP)((\phi A)l,\id_{MA}) \alpha_{A,A,M} x, (UQ)(\id_{MA},r) \beta_{A,A,M} y\rangle && \text{(by the coend relations again)} \\
%&= \langle (UP)(lrl,\id_{MA}) \alpha_{A,A,M} x, (UQ)(\id_{MA},r) \beta_{A,A,M} y\rangle && \text{(as $lr = \phi A$ again)} \\
%&= \langle (UP)(l,\id_{MA}) \alpha_{A,A,M} x, (UQ)(\id_{MA},r) \beta_{A,A,M} y\rangle && \text{(as $rl = \id_S$)} \\
%&= (U(P \odot Q))(l,r) \langle \alpha_{A,A,M} x, \beta_{A,A,M} y \rangle && \text{(by definition of $P \odot Q$)} \\
%&= (U(P \odot Q))(l,r) (\alpha_{A,A,M} \odot \beta_{A,A,M}) \langle x, y \rangle && \text{(by definition of $\alpha \odot \beta$)} \\
%&= \tilde{p}_{P \odot Q} \langle x, y \rangle
%And for the second:
%\tilde{p}_I(\id_A) = (UI)(l,r) (M \id_A) = r \id_{MA} l = rl = \id_S
%as required.
%%\begin{proposition}
